The Symmetry of Partner Modelling

### 2016-02-29 mutual-modeling.tex
Improved a bit the quality of 2 figures
Last changes to the synthesis

\begin{figure*}
        \centering
        \includegraphics[width=\textwidth]{image4.png}
        \caption{Screenshot of the {\sc SpaceMiners} game}
        \label{study1:spaceminer}
\end{figure*}

@@

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.6\textwidth}
            \includegraphics[width=\linewidth]{study5-conceptmap.png}
        \end{subfigure} \\
        \begin{subfigure}{.7\textwidth}
            \includegraphics[width=\linewidth]{study5-kat.png}
        \end{subfigure}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

@@

We found a symmetry of the mutual models on 4 studies out of 5. In study 4, this
only applies to the control group (having read the same text before), which
corresponds to the situation of the 3 first studies. Evidencing this symmetry
constitutes \textit{per se} an interesting result as we are not aware of earlier
studies that have established this relationship.
Still, the symmetry alone does not allow to discriminate $\mathcal{H}_{2}$ from
$\mathcal{H}_{3}$. The second hypothesis (\ie mutual modelling is primarily
supported by the capacity of the modellee to help the modellers to form a
correct image of herself), is questioned by study 3 where \C could be modelled
accurately by \A and not accurately by \B. The same hypothesis is however
supported by the results of study 2, and indirectly by study 4.

@@

We acknowledge that the difference between these hypotheses is rather
theoretical since the  process of modelling one's partner $\mathcal{H}_{1}$ and
the process of helping one's partner to model oneself $\mathcal{H}_{2}$ are
mediated by verbal interactions in the team. One could hardly imagine someone
managing a very accurate modelling despite the very bad quality of interactions.
There is a bidirectional causal link between the accuracy of mutual modelling
and the quality of the interactions. This rather artificial distinction does
however participate in the more fundamental discussion of the role of individual
and social mechanisms in human cognition. In the field of social cognition, it
is a commonplace to state that ``the whole is greater than the sum of the
parts''.  This refers to the emergence of team properties than cannot be reduced
to the set of individual contributions. Our paper illustrates this emergence by
showing the symmetry of mutual modelling.

@@

Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do
not imply we have a mechanical view of modelling. The notation is just a language for
referring to different components of modelling each other in teams.  This simple
notation emerged from the need to compare five mutual modelling studies. It
could however pave the road for elaborating a systematic agenda for research on
mutual modelling.

>>>>

\begin{figure*}
        \centering
        \includegraphics[width=\textwidth]{study1}
        \caption{Screenshot of the {\sc SpaceMiners} game}
        \label{study1:spaceminer}
\end{figure*}

@@

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.6\textwidth}
            \includegraphics[width=\linewidth]{study5-conceptmap.pdf}
        \end{subfigure} \\
        \begin{subfigure}{.7\textwidth}
            \includegraphics[width=\linewidth]{study5-kat.png}
        \end{subfigure}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

@@

In the introduction, we mentioned the controversy around the cognitive depth of
dialogues: do they build upon shallow syntactic alignment or do they entail
deeper semantic processes, which would require modelling of the partner. Our
studies rather support the second viewpoint since the quality of modelling
emerged as an intermediate variable, sensitive to some independent variables
(like the presence of awareness tools or media richness), and predictive of
dependent variables, such as task performance. We do not claim that this deeper
semantic modelling is always present in everyday conversations, but rather that
it \emph{was} indeed present in the collaborative tasks designed for our
experiments (which are representative of the tasks assigned to learners in CSCL
environments).

Secondarily, we questioned whether the accuracy of the partner's model depends on
the cognitive skills of each partner or instead results from the quality of
the interactions among them. We respectively refer to these hypothesis as
$\mathcal{H}_1$ and $\mathcal{H}_3$. Our rationale was that a symmetry of mutual
modelling (correlation between the accuracy of each other's model) would favor
$\mathcal{H}_3$ over $\mathcal{H}_1$. We also formulated $\mathcal{H}_2$, by
which a learner \A who is good at modelling \B would also help \B to repair his
inaccurate representations of \A.

We found a symmetry of the mutual models on 4 studies out of 5. In study 4, this
only applies to the control group (having read the same text before), which
corresponds to the situation of the 3 first studies. Evidencing this symmetry
constitutes \textit{per se} an interesting result as we are not aware of earlier
studies that have established this relationship.  Still, the symmetry alone does
not allow to discriminate $\mathcal{H}_{2}$ from $\mathcal{H}_{3}$. The second
hypothesis is questioned by study 3 where \C could be modelled accurately by \A
and not accurately by \B. The same hypothesis is however supported by the
results of study 2, and indirectly by study 4.

@@

We acknowledge that the difference between these hypotheses is rather
theoretical since the  process of modelling one's partner $\mathcal{H}_{1}$ and
the process of helping one's partner to model oneself $\mathcal{H}_{2}$ are
mediated by verbal interactions in the team. One could hardly imagine someone
managing a very accurate modelling despite the very bad quality of interactions.
There is a bidirectional causal link between the accuracy of mutual modelling
and the quality of the interactions. This rather artificial distinction does
however participate in the more fundamental discussion of the role of individual
and social mechanisms in human cognition. In the field of social cognition, it
is a commonplace to state that ``the whole is greater than the sum of the
parts''.  This refers to the emergence of team properties than cannot be reduced
to the set of individual contributions. Our paper illustrates this emergence by
showing the symmetry of mutual modelling. In a nutshell, yes there is a
non-negligible component of mutuality in modelling one's partner.

@@

These limitations and our difficulty to provide clear-cut conclusions comes from
the fact that this series of experiments was not planned a priori. This paper
relies on the post-hoc comparison of experiments conducted across various
contexts. The overall conclusion is that this research question would deserve a
specific research agenda for 3 reasons: (1) it raises fundamental theoretical
dilemma on the social nature of cognition, (2) it raises methodological
challenges and (3) it could provide empirical grounds to design decisions for
CSCL environments. The eye tracking tools we developed and the notation we used
for referring to different components of mutual modelling could pave the road
for elaborating a systematic agenda for research on mutual modelling. Let us
repeat that the expressions we used, such as \gmodel{A}{B}, do not imply we have a
mechanical view of modelling, but were simply useful ways to talk about mutual
modelling.
